### 2015-05-04 papers/histogram/thesis/thesis.tex
modifications to a derivation for walker current

The down-going walker flux $j\p{E}$ is defined in terms of the
diffusivity $\alpha\p{E}$ by
\begin{align}
  j=\alpha\f{dn_+}{dE}.
  \label{eq:walker_current_definition}
\end{align}
Computing
\begin{align}
  n\f{df}{dE}=n\sp{\f1{n}\f{dn_+}{dE}-\f{n_+}{n^2}\f{dn}{dE}}
  =\f{dn_+}{dE}-f\f{dn}{dE},
  \label{eq:walker_intermediate}
\end{align}
we can say that
\begin{align}
  \f{dn_+}{dE}=n\f{df}{dE}+f\f{dn}{dE}.
  \label{eq:walker_gradient}
\end{align}
Substituting (\ref{eq:walker_gradient}) into
(\ref{eq:walker_current_definition}) yields
\begin{align}
  j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
  \label{eq:walker_current_full}
\end{align}
If the energy histogram does vary strongly with energy, e.g. because
we are simulating with flat histogram weights, then $dn/dE\approx0$,
which means
\begin{align}
  j=\alpha n\f{df}{dE}.
  \label{eq:walker_current}
\end{align}
To formalize the approximation made between
(\ref{eq:walker_current_full}) and (\ref{eq:walker_current}), we can
observe that
\begin{align}
  \f{f~dn/dE}{n~df/dE}
  =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}=\p{\f1f\f{dn_+}{dn}-1}^{-1}
  =\p{\f{d\ln n_+}{d\ln n}-1}^{-1}.
  \label{eq:walker_current_intermediate_ratio}
\end{align}
For (\ref{eq:walker_current}) to be a valid, we want the ratio in
(\ref{eq:walker_current_intermediate_ratio}) to vanish. If the walker
density $n$ varies weakly with energy, e.g. because we are simulating
with flat histogram weights, then $\ln n$ varies even weaker
still. The down-going walker density $n_+$, meanwhile must run from
$n\p{E_-}=0$ to $n_+\p{E_+}=n\p{E_+}$ meaning $\ln n_+$ runs from $\ln
n\p{E_-}\to-\infty$ to a finite positive value $\ln n_+\p{E_+}=\ln
n\p{E_+}$. It is therefore reasonable to assume that $\abs{d\ln
  n_+/d\ln n}\gg 1$, which makes the ratio in
\ref{eq:walker_current_intermediate_ratio} vanish.

We can rearrange (\ref{eq:walker_current}) and integrate from the low
energy state $E_-$, at which $f\p{E_-}=0$, to the high energy state
$E_+$, at which $f\p{E_+}=1$, to say
\begin{align}
  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_setup}
\end{align}
At equilibrium, the walker flux $j\p{E}$ must be constant with respect
to energy. If $j\p{E}$ is not constant with respect to energy, some
energies would have either a net influx or net outflow of walkers,
which would mean that the ensemble is not at equilibrium. We can
therefore evaluate one of the integrals in
(\ref{eq:walker_current_setup}) to get
\begin{align}
  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral}
\end{align}

The optimized ensemble method minimizes the integral in
(\ref{eq:walker_current_integral}), thereby maximizing the down-going
walker flux $j$, by varying the walker density $n$ with the constraint
that $n$ must remain normalized. With the additional assumption that
$\alpha\p{E}$ does not strongly depend on the weights $w\p{E}$, the
authors of \cite{optimized_ensemble} find that the optimal walker
density $n$ is
\begin{align}
  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
  \label{eq:optimal_walker_density}
\end{align}
where $\lambda$ is the Lagrange multiplier used in the minimization
process, and here acts merely as a normalization factor.

>>>>

The down-going walker current $j\p{E}$ is defined in terms of the
diffusivity $\alpha\p{E}$ by
\begin{align}
  j=\alpha\f{dn_+}{dE}.
  \label{eq:walker_current_definition}
\end{align}
Computing
\begin{align}
  n\f{df}{dE}=n\sp{\f1{n}\f{dn_+}{dE}-\f{n_+}{n^2}\f{dn}{dE}}
  =\f{dn_+}{dE}-f\f{dn}{dE},
  \label{eq:walker_intermediate}
\end{align}
we can say that
\begin{align}
  \f{dn_+}{dE}=n\f{df}{dE}+f\f{dn}{dE}.
  \label{eq:walker_gradient}
\end{align}
Substituting (\ref{eq:walker_gradient}) into
(\ref{eq:walker_current_definition}) yields
\begin{align}
  j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
  \label{eq:walker_current_intermediate}
\end{align}
We can observe that
\begin{align}
  \f{f~dn/dE}{n~df/dE} =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}
  =\p{\f1f\f{dn_+}{dn}-1}^{-1}.
  \label{eq:walker_current_approx_ratio_setup}
\end{align}
Furthermore,
\begin{align}
  \p{\f1f\f{dn_+}{dn}-1}^{-1}
  =\p{\f1f\f{d\p{fn}}{dn}-1}^{-1}
  =\p{\f nf\f{df}{dn}+\f{dn}{dn}-1}^{-1}
  =\p{\f nf\f{df}{dn}}^{-1}=\f nf\f{dn}{df}=\f{d\ln n}{d\ln f},
  \label{eq:walker_current_approx_intermediate}
\end{align}
which means
\begin{align}
  \f{f~dn/dE}{n~df/dE}=\f{d\ln n}{d\ln f}.
  \label{eq:walker_current_approx_ratio}
\end{align}
The walker current is therefore
\begin{align}
  j=\alpha n\f{df}{dE}\p{1+\f{d\ln n}{d\ln f}}.
  \label{eq:walker_current_full}
\end{align}
We know that $f$ runs from $f\p{E_-}=0$ to $f\p{E_+}=1$ meaning $\ln
f$ runs from $\ln f\p{E_-}\to-\infty$ to $\ln f\p{E_+}=0$. If the
walker density $n$ varies weakly with energy, e.g. because we are
simulating with flat histogram weights, then $\ln n$ varies even
weaker still. We may therefore reasonably to assume that $\abs{d\ln
  n/d\ln f}\ll 1$ and express
\begin{align}
  j=\alpha n\f{df}{dE}.
  \label{eq:walker_current}
\end{align}
We can rearrange (\ref{eq:walker_current}) and integrate over the
energy range to say
\begin{align}
  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral_setup}
\end{align}
At equilibrium, the walker current $j\p{E}$ must be constant with
respect to energy. If $j\p{E}$ is not constant with respect to energy,
some energies would have either a net incurrent or net outflow of
walkers, which would mean that the ensemble is not at equilibrium. We
can therefore evaluate one of the integrals in
(\ref{eq:walker_current_integral_setup}) to get
\begin{align}
  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral}
\end{align}

The optimized ensemble method minimizes the integral in
(\ref{eq:walker_current_integral}), thereby maximizing the down-going
walker current $j$, by varying the walker density $n$ with the
constraint that $n$ must remain normalized. With the additional
assumption that $\alpha\p{E}$ does not strongly depend on the weights
$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
optimal walker density $n$ is
\begin{align}
  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
  \label{eq:optimal_walker_density}
\end{align}
where $\lambda$ is the Lagrange multiplier used in the minimization
process, and here acts merely as a normalization factor.
